ShrinkNets: Learning Network Size while Training

### 2018-02-09 paper/ICML/introduction.tex

Neural networks are increasingly being deployed in a wide variety  of
applications  and on diverse hardware architectures ranging from laptops to
phones to embedded sensors. This wide variety of deployment settings means
that inference time and model size are becoming as important as  prediction
quality when assessing model performance.  However, these three dimensions of
performance (model quality, inference time and model size)  are largely
optimized separately today, often with suboptimal results.

Of course, the problem of finding compressed or small networks is not new.
Existing techniques to make a pre-trained neural network smaller can be
categorized in two approaches: (1) quantization \cite{quant} and code
compilation~\cite{ma2016compilation}, techniques that can be applied blindly to
any network, and (2) techniches which analyze the structure of the network and
systematically prune connections or neurons~\cite{han2015deepcompression,Cun}
based on some loss function. While the first category is always useful, but only
has limited impact on the network size, the second category can reduce the model
size much more but has several drawbacks:   First, those techniques often
negatively impact model quality.  Second, they can (surprisingly) negatively
impact inference time as they transform dense matrix operations into sparse
ones, which can be substantially slower to execute on modern hardware
\cite{han2015deepcompression}, especially GPUs which do not efficiently support
sparse linear algebra.  Third, these techniques generally start by optimizing a
particular architecture for prediction performance, and then, as a post-
processing step, applying compression  to generate a smaller model that meets
the resource constraints of the deployment setting.  Because the network
architecture is essentially fixed during this post-processing,   model
architectures that work better in small settings may be missed -- this is
especially true in large networks like many-layered CNNs,  where it is
infeasible to try explore even a small fraction of possible network
configurations.

@@

In contrast, in this paper we present a new method to simultaneously optimize
 network size and model performance. The key idea is to learn the right
network size at the same time that we optimize for prediction performance for a
specific task. Our approach, called {\it ShrinkNets}, starts with an {\it
over-sized} network, and dynamically shrinks it by eliminating unimportant
neurons---those that do not contribute to prediction performance---during
training time. To do this, ShrinkNets must detect the unimportant neurons, and
remove them from the network. The
removal of entire neurons and not connections is crucial, because it leads to
dense networks, which in turn leads to better inference performance. To support the
detection and removal of unimportant neurons we introduce a new layer, called
\textbf{Switch layer}. Introducing this new layer simply requires us to
add a term to the
loss function to optimize the weights of this layer during training. ShrinkNets has
two main benefits. First, it explores the architecture of models that are both
small and perform well, rather than starting with a high-performance model and
making it small.  It does this using a single new hyperparameter that effectively 
models the target network size.  This allows us to efficiently generate a family of smaller and
accurate models without an exhaustive and expensive hyperparameter search over
the number of neurons in each layer.  Second, in contrast to existing neural
network compression techniques~\cite{Aghasi2016,han2015deepcompression}, our
approach results in models that are not only small, but where the layers are
dense, which means that inference time is also improved, e.g., on GPUs.

>>>>

Neural networks are being deployed in a wide variety of
applications and on diverse hardware architectures ranging from laptops to
phones to embedded sensors. This wide variety of deployment settings means
that inference time and model size are becoming as important as  prediction
accuracy when assessing model quality.  However, these three dimensions of
performance (prediction accuracy, inference time and model size) are largely
optimized independently, often with suboptimal results.

Of course, the problem of finding compressed or small networks is not new.
Existing techniques to make a pre-trained neural network smaller can be
categorized in two approaches: (1) quantization \cite{Jouppi:2017:IPA:3079856.3080246} and code
compilation, techniques that can be applied blindly to
any network, and (2) techniques which analyze the structure of the network and
systematically prune connections or neurons~\cite{han2015deepcompression,Cun}. 
While the first category is always useful, but only
has limited impact on the network size, the second category can reduce the model
size much more but has several drawbacks:   First, those techniques often
negatively impact model quality.  Second, they can (surprisingly) negatively
impact inference time as they transform dense matrix operations into sparse
ones, which can be substantially slower to execute on GPUs which do not support
efficiently sparse linear algebra\cite{han2015deepcompression}. 
Third, these techniques generally start by optimizing a
particular architecture for prediction performance, and then, as a post-
processing step, applying compression  to generate a smaller model that meets
the resource constraints of the deployment setting.  Because the network
architecture is essentially fixed during this post-processing,   model
architectures that work better in small settings may be missed -- this is
especially true in large networks like many-layered CNNs,  where it is
infeasible to try explore even a small fraction of possible network
configurations.

@@

In contrast, in this paper we present a new method to simultaneously optimize
 network size and model performance. The key idea is to learn the right
network size at the same time that we optimize for prediction performance. 
Our approach, called \shrink, starts with an {\it
over-sized} network, and dynamically shrinks it by eliminating unimportant
neurons---those that do not contribute to prediction performance---during
training time. To do this, \shrink must detect the unimportant neurons, and
remove them from the network. The
removal of entire neurons (vs. only connections) is crucial, because it leads to
dense networks, which in turn leads to better inference performance. To support the
detection and removal of unimportant neurons we introduce a new layer, called
\swl. Introducing this new layer simply requires us to
add a term to the
loss function to optimize the parameters of this layer during training. 
\shrink has
two main benefits. First, it explores the architecture of models that are both
small and perform well, rather than starting with a high-performing model and
making it small.  \shrink accomplishes this goal by using a single new 
hyperparameter that effectively models the target network size.  
% This allows us to efficiently generate a family of smaller and
% accurate models without requiring an exhaustive and expensive hyperparameter search over
% the number of neurons in each layer.  
Second, in contrast to existing neural
network compression techniques~\cite{Aghasi2016,han2015deepcompression}, our
approach results in models that are not only small, but where the weight matrices
are dense, leading to better inference time.
